%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

I first thank my extraordinary supervisor, Dr.~Karon MacLean, for all of her advice, direction, and support throughout my graduate studies.
I am extremely fortunate to receive her mentorship.

I am very grateful to my supervisory and examination committees. Dr.~Ronald Garcia and Dr.~Michiel van de Panne always made time for insightful, thought-provoking, and helpful discussions that have thoroughly enriched this work.
Dr.~Antony Hodgson, Dr.~Eric Meyers, Dr.~Luanne Freund, and my external examiner Dr.~Ian Oakley asked challenging, valuable questions that led to several key improvements of this dissertation.
I also thank the other faculty who provided feedback over the years: Dr.~Kellogg Booth, Dr.~Joanna McGrenere, and Dr.~Tamara Munzner.

I have immense appreciation for the friends and colleagues that have made the last four years an amazing experience.
Each of you left an indelible mark.
At UBC, Dr.~Idin Karuei, Hasti Seifi, Laura Cang, Paul Bucci,  David Marino, Soheil Kianzad, Matthew Chun, Salma Kashani, Gordon Minaker, Brenna Li,  Ben Clark, Dilan Ustek, Meghana Venkatswany, Dr.~Matthew Brehmer, Jessica Dawson, Anamaria Crisani, Kamyar Ardekani, Antoine Ponsard, Derek Cormier, Francisco Escalona;
at Disney Research, Dr.~Ali Israr, Siyan Zhao; 
at Stanford University, Dr.~Allison Okamura, Melisa Orta Martinez, Richard Davis;
at McGill University, Jan Anlauff, Colin Gallacher, Jeff Blum;
and beyond this list, there are many more.

I also greatly appreciate my funding sources for supporting this work.
I thank the Natural Sciences and Engineering Research Council of Canada (NSERC), the National Science Foundation (NSF), Disney Research Pittsburgh, and the University of British Columbia.

Finally, I am forever grateful to my family for their undying love and support.
%Dr.~Kevin Schneider, Dr.~Karen Wood, Samuel Schneider, Anna Schneider, and Elyse Clancy all had a hand in my progress.